---
output: 
  bookdown::github_document2:
    number_sections: false
bibliography: "docs/ecgref.bib"
biblio-style: "apalike"
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = .618
  )
options(digits = 3)
```

# Electrocardiogram Experiment

## About the project

This is a project performed in [SKKU **Design and Analysis of Experiments**](https://www.kwonsanglee.com) [@sta5031].

- Conduct a real randomized experiment, collect data, and analyze them.
- Present a non-statistical paper using an (quasi-)experiment in the field outside statistics and critize them.
- Bring my own research that is related to an (quasi-)experiment.

I chose real experimental design, the first one.

## Goal of the Experiment

- **Does caffeine affect electrocardiogram (ECG)?**
- Caffeine: drinking capsule coffee 40 ml
- ECG result: average heart rate

## Design

We implement Latin square design [@Fisher:1935uc]. For example,

<table>
  <caption>Reduced latin square</caption>
  <thead>
    <tr>
      <td colspan = 2></td>
      <th colspan=5>Drinking speed</th>
    </tr>
  </thead>    
  <tbody>
    <tr>
      <td colspan = 2></td>
      <td>1</td>
      <td>2</td>
      <td>3</td>
      <td>4</td>
    </tr>
    <tr>
      <th rowspan=4>Coffee to water ratio</th>
      <td>1</td>
      <td>A</td>
      <td>B</td>
      <td>C</td>
      <td>D</td>
    </tr>
    <tr>
      <td>2</td>
      <td>B</td>
      <td>C</td>
      <td>D</td>
      <td>A</td>
    </tr>
    <tr>
      <td>3</td>
      <td>C</td>
      <td>D</td>
      <td>A</td>
      <td>B</td>
    </tr>
    <tr>
      <td>4</td>
      <td>D</td>
      <td>A</td>
      <td>B</td>
      <td>C</td>
    </tr>
  </tbody>
</table>

- A, B, C, D once in every row,
- once in every column

### About factors
- Latin square: 4 by 4
- 2 blocking factors
    - **Row**: Coffee (40 ml) to water ratio
        1. 1:0 (Espresso)
        2. 1:2.5 (Water 100 ml)
        3. 1:5 (Water 200 ml)
        4. 1:7.5 (Water 300 ml)
    - **Column**: Drinking speed
        1. <=5 sec
        2. 5-15 sec
        3. 15-30 sec
        4. 30< sec
- Interesting factor: [Intake of caffeine](https://www.reddit.com/r/nespresso/comments/id31r5/i_recieved_the_caffiene_content_numbers_for/) from [Starbucks by Nespresso](https://athome.starbucks.com/coffees-by-format/nespresso-original/)
    1. House blend: 74.5 mg
    2. Sumatra: 54.5 mg
    3. Decaf espresso roast: 3 mg
    4. None or water: 0 mg
- Randomly allocate these treatments `1`, `2`, `3`, and `4` to A, B, C, and D
    - assign these to above table
    - random treatment assignment

```{r}
set.seed(1)
sample(LETTERS[1:4])
```

### Output
- Measure ECG using *Apple watch Series 4*: See [https://support.apple.com/en-us/HT208955](https://support.apple.com/en-us/HT208955)
- Output: **Average Heart rate (in BPM)** difference

### Controlling other variables
- Drink coffee every morning (between 8:30 a.m. and 9:00 a.m. KST), after eating a piece of bread
- When Measuring ECG,
    - sit at my desk
    - rest arms on a my desk
- Use same strip for apple watch: [Nike sport band](https://www.apple.com/shop/product/MX8C2AM/A/40mm-anthracite-black-nike-sport-band-regular?fnode=5e9ad1340eb02decfee1689be9360555f2f276ad270a672413266cfba01ad7b0e20a1c634dbd66eaec20c01170cf533573070d71c910b376e339037f157174b7e6f45e144d64e052e5274d1069eb67b4) of same fit
- Nespresso machine: [Pixie C61](https://www.nespresso.com/kr/en/order/machines/original/pixie-electric-red-coffee-machine) in my home

---
