---
output: github_document
bibliography: "docs/ecgref.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.asp = .618
  )
options(digits = 3)
```

# Electrocardiogram Experiment

## Goal

- **Does caffeine affect electrocardiogram (ECG)?**
- Caffeine: drinking capsule coffee 40 ml

## Design

We use Latin square design [@Box:2005ug].

### About factors
- Latin square: 4 by 4
- 2 blocks components
    - Coffee (40 ml) to water ratio
        - 1:0 (Espresso)
        - 1:2
        - 1:8
        - 1:15
    - Drinking speed
        - <=5 sec
        - 5-20 sec
        - 20-50 sec
        - 50< sec
- Interesting factor: [Intake of caffeine](https://www.reddit.com/r/nespresso/comments/id31r5/i_recieved_the_caffiene_content_numbers_for/) from [Starbucks by Nespresso](https://athome.starbucks.com/coffees-by-format/nespresso-original/)
    1. House blend: 74.5 mg
    2. Sumatra: 54.5 mg
    3. Decaf espresso roast: 3 mg
    4. None or water: 0 mg
- Randomly allocate these treatments to A, B, C, and D: inside Latin square

```{r}
set.seed(1)
sample(LETTERS[1:4])
```

### Output
- Measure ECG using *Apple watch Series 4*: See [https://support.apple.com/en-us/HT208955](https://support.apple.com/en-us/HT208955)
- Output: **Average Heart rate (in BPM)**

### Controlling other variables
- Drink coffee every morning
- Before breakfast
- Measure ECG in the same pose, same location
- Water temperature

---
