---
title: "Electrocardiogram Experiment"
subtitle: "using latin square design"
author: |
  | [Young Geun Kim](mailto:dudrms33@g.skku.edu)
  | [ygeunkim.github.io](https://ygeunkim.github.io)
institute: "2019711358, [Department of Statistics](https://stat.skku.edu/stat/index.jsp)"
date: "`r format(Sys.time(), '%d %b, %Y')`"
bibliography: "../docs/ecgref.bib"
# biblio-style: "apalike"
link-citations: yes
nocite: |
  @sta5031
output: 
  beamer_presentation:
    toc: yes
    slide_level: 2
    theme: "Darmstadt"
    colortheme: "beaver"
    fonttheme: "professionalfonts"
    citation_package: biblatex
    includes:
      in_header: "../docs/preamble.tex"
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(input = inputFile, encoding = encoding, output_dir = "../static/slides")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  # fig.pos = "H",
  fig.asp = .618
  )
knitr::knit_hooks$set(
  document = function(x) {
    sub("\\usepackage[]{color}", "\\usepackage{xcolor}", x, fixed = TRUE)
  }
)
options(digits = 3)
options(kableExtra.latex.load_packages = FALSE)
is_latex <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "latex"
is_beamer <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "beamer"
```

```{r pkgs, message=FALSE, echo=FALSE}
# tidyverse family---------------------
library(tidyverse)
# kable--------------------------------
library(knitr)
library(kableExtra)
# set seed for report -----------------
set.seed(1)
```

# Introduction

## Electrocardiogram Experiment

## Latin Square Design


<!-- details of the design -->

# Design and the Data

## Block Components

## Factor

## Randomization

## Output

## Controlling the Other Variables

### Coffee

- Drink coffee every morning
- before eating breakfast
- Nespresso machine: [Pixie C61](https://www.nespresso.com/kr/en/order/machines/original/pixie-electric-red-coffee-machine) in my home

### Measure

- Sitting at the table
- Rest my arms on the table

### Apple watch

- Use the same strip
    - [Modern buckle](https://www.apple.com/shop/product/MY662AM/A/40mm-scarlet-modern-buckle-small?fnode=a752f02930ae068b20b0bd9682d5cc42653383b01f5d691eb1ced059ce28de220bacf6009213e93e9c00295f48285a7ed49f95a0721d433ea8c006ff7772eb3d0f0dab89c6c9538ff9063124f30717a6)
    - since its length is fixed

<!-- analysis -->

# Analysis

## ANOVA

<!-- references -->

# Related Contents

## Project Repo

My Github project for this experiment:

[https://github.com/ygeunkim/ecg-experiment](https://github.com/ygeunkim/ecg-experiment)

