---
title: "Supplementary Materials"
subtitle: "Datasets"
author: |
  | Young Geun Kim
  | [ygeunkim.github.io](https://ygeunkim.github.io)
date: "`r format(Sys.time(), '%d %b, %Y')`"
bibliography: "../docs/ecgref.bib"
biblio-style: "apalike"
link-citations: yes
output: 
  bookdown::pdf_document2:
    toc: yes
    latex_engine: lualatex
    includes:
      in_header: "../docs/preamble.tex"
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(input = inputFile, encoding = encoding, output_dir = "../static/appendix")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  out.width = "70%",
  fig.align = "center",
  fig.width = 6,
  fig.pos = "H",
  fig.asp = .618
  )
knitr::knit_hooks$set(
  document = function(x) {
    sub("\\usepackage[]{color}", "\\usepackage{xcolor}", x, fixed = TRUE)
  }
)
options(digits = 3)
options(knitr.kable.NA = "")
options(kableExtra.latex.load_packages = FALSE)
is_html <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "html"
is_latex <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "latex"
kable_format <- ifelse(!is_html, "latex", "html")
kable_format <- ifelse(length(kable_format) == 0, "html", kable_format) # for excecutable
```

```{r pkgs, message=FALSE, echo=FALSE}
# tidyverse family---------------------
library(tidyverse)
# kable--------------------------------
library(knitr)
library(kableExtra)
```

# Data

We provide experiment data here.
In the slide, some part of the tables were skipped due to its size.

## Experiment Data

```{r readdata, message=FALSE, echo=FALSE}
# Import---------------------------------
ecg <- read_csv("../data/raw/ecg.csv")
ecg_diff <- read_csv("../data/processed/ecg-diff.csv")
```

The following is the raw dataset:

```{r}
ecg
```

Table \@ref(tab:diffdata) is LSD format based on $y_{rc} \defn y_{rc}^{post} - y_{rc}^{pre}$.

```{r diffdata, echo=FALSE}
# Print----------------------------------
ecg_diff %>% 
  select(-id, -date) %>% 
  mutate(
    water = case_when(
      water == 1 ~ "0 ml",
      water == 2 ~ "100 ml",
      water == 3 ~ "200 ml",
      water == 4 ~ "300 ml"
    ),
    speed = case_when(
      speed == 1 ~ "<=5",
      speed == 2 ~ "5-15",
      speed == 3 ~ "15-30",
      speed == 4 ~ "30<"
    ),
    coffee = case_when(
      coffee == 1 ~ "HB",
      coffee == 2 ~ "S",
      coffee == 3 ~ "D",
      coffee == 4 ~ "W"
    )
  ) %>% 
  mutate(
    diff = cell_spec(
      diff, 
      color = spec_color(diff, end = .9, direction = -1),
      font_size = spec_font_size(diff, begin = 8)
    )
  ) %>% 
  unite(coffee, diff, col = "output", sep = ", ") %>% 
  group_by(water) %>% 
  pivot_wider(names_from = speed, values_from = output) %>% 
  kable(
    format = kable_format,
    escape = FALSE,
    caption = "Experiment Data"
  ) %>% 
  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position")) %>% 
  add_header_above(
    header = c(" " = 1, "Drinking Speed" = 4),
    escape = FALSE
  ) %>% 
  footnote(
    number = c(
      "Caffeine: HB > S > D > W",
      "Numbers indicate the difference after and before taking coffee"
    )
  )
```

## Science Table

To test sharp null of no effect

\begin{equation}
  H_0: y_{rc}(1) = y_{rc}(2) = y_{rc}(3) = y_{rc}(4)
  (\#eq:sharpnoeff)
\end{equation}

we build the science table in Table \@ref(tab:posharp).

```{r posharp, message=FALSE, echo=FALSE}
# science table-----------------------------------------------------------
ecg_po <- read_csv("../data/processed/science.csv")
ecg_po %>% 
  # slice(1:4, 13:16) %>%
  kable(
    format = kable_format,
    escape = FALSE,
    caption = "Observed Values of the Science Table"
  ) %>% 
  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position")) %>% 
  add_header_above(
    header = c(" " = 4, "Observed $y_{rc}(k)$" = 4),
    escape = FALSE
  ) %>% 
  pack_rows(
    "Row 1 (Water 0 ml)",
    1, 4
  ) %>% 
  pack_rows(
    "Row 2 (Water 100 ml)",
    5, 8
  ) %>% 
  pack_rows(
    "Row 3 (Water 200 ml)",
    9, 12
  ) %>% 
  pack_rows(
    "Row 4 (Water 300 ml)",
    13, 16
  )
```

Now we impute impute the missing $Y_{rc}(k)$ under the sharp null. See Table \@ref(tab:tablesharp).

```{r tablesharp, message=FALSE, echo=FALSE}
# imputing-------------------------------------------------------------------
ecg_sharp <- read_csv("../data/processed/science-sharp.csv")
ecg_sharp %>% 
  # slice(1:4, 13:16) %>%
  kable(
    format = kable_format,
    escape = FALSE,
    caption = "Imputed Outcomes under the Sharp Null",
    longtable = TRUE
  ) %>% 
  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position")) %>% 
  add_header_above(
    header = c(" " = 4, "Observed $y_{rc}(k)$" = 4),
    escape = FALSE
  ) %>% 
  pack_rows(
    "Row 1 (Water 0 ml)",
    1, 4
  ) %>% 
  pack_rows(
    "Row 2 (Water 100 ml)",
    5, 8
  ) %>% 
  pack_rows(
    "Row 3 (Water 200 ml)",
    9, 12
  ) %>% 
  pack_rows(
    "Row 4 (Water 300 ml)",
    13, 16
  )
```

\newpage

## After Presentation

In general, heart rate variable is used as log scale, i.e. here, log-return.

```{r readlog, message=FALSE, echo=FALSE}
ecg_log <- read_csv("../data/processed/ecg-logreturn.csv")
sharp_log <- read_csv("../data/processed/science-sharp-log.csv")
```

See Table \@ref(tab:logdata).

```{r logdata, echo=FALSE}
# Print----------------------------------
ecg_log %>% 
  select(-id, -date) %>% 
  mutate(
    water = case_when(
      water == 1 ~ "0 ml",
      water == 2 ~ "100 ml",
      water == 3 ~ "200 ml",
      water == 4 ~ "300 ml"
    ),
    speed = case_when(
      speed == 1 ~ "<=5",
      speed == 2 ~ "5-15",
      speed == 3 ~ "15-30",
      speed == 4 ~ "30<"
    ),
    coffee = case_when(
      coffee == 1 ~ "HB",
      coffee == 2 ~ "S",
      coffee == 3 ~ "D",
      coffee == 4 ~ "W"
    )
  ) %>% 
  mutate(
    diff = cell_spec(
      diff %>% round(3), 
      color = spec_color(diff, end = .9, direction = -1),
      font_size = spec_font_size(diff, begin = 8)
    )
  ) %>% 
  unite(coffee, diff, col = "output", sep = ", ") %>% 
  group_by(water) %>% 
  pivot_wider(names_from = speed, values_from = output) %>% 
  kable(
    format = kable_format,
    escape = FALSE,
    caption = "Experiment Data - log-return for the average heart rate"
  ) %>% 
  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position")) %>% 
  add_header_above(
    header = c(" " = 1, "Drinking Speed" = 4),
    escape = FALSE
  ) %>% 
  footnote(
    number = c(
      "Caffeine: HB > S > D > W",
      "Numbers indicate the difference after and before taking coffee"
    )
  )
```

Table \@ref(tab:logsharp) is the imputed science table.

```{r logsharp, message=FALSE, echo=FALSE}
# imputing-------------------------------------------------------------------
sharp_log %>% 
  kable(
    format = kable_format,
    escape = FALSE,
    caption = "Imputed Outcomes under the Sharp Null - log-return",
    longtable = TRUE
  ) %>% 
  kable_styling(full_width = FALSE, latex_options = c("striped", "HOLD_position")) %>% 
  add_header_above(
    header = c(" " = 4, "Observed $y_{rc}(k)$" = 4),
    escape = FALSE
  ) %>% 
  pack_rows(
    "Row 1 (Water 0 ml)",
    1, 4
  ) %>% 
  pack_rows(
    "Row 2 (Water 100 ml)",
    5, 8
  ) %>% 
  pack_rows(
    "Row 3 (Water 200 ml)",
    9, 12
  ) %>% 
  pack_rows(
    "Row 4 (Water 300 ml)",
    13, 16
  )
```

\newpage

# ECG Results

From ECG app, we get the graph as in Figure \@ref(fig:resplot).

```{r, include=FALSE}
fig_subcap <- c(
  "(1,1),House blend", "(1,2),Water", "(1,3),Sumatra", "(1,4),Decaf",
  "(2,1),Water", "(2,2),Sumatra", "(2,3),Decaf", "(2,4),House blend",
  "(3,1),Sumatra", "(3,2),Decaf", "(3,3),House blend", "(3,4),Water",
  "(4,1),Decaf", "(4,2),House blend", "(4,3),Water", "(4,4),Sumatra"
)
fig_subcap <- str_c(fig_subcap, str_c(ecg$post, " BPM"), sep = ": ")
```

```{r resplot, echo=FALSE, fig.cap="Electrocardiogram after drinking Coffee", fig.subcap = fig_subcap, fig.asp=1, fig.ncol = 4, out.width="25%"}
# (1, 1)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-22.pdf")
# (1, 2)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-23.pdf")
# (1, 3)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-24.pdf")
# (1, 4)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-25.pdf")
# (2, 1)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-26.pdf")
# (2, 2)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-27.pdf")
# (2, 3)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-28.pdf")
# (2, 3)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-29.pdf")
# (3, 1)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-04-30.pdf")
# (3, 2)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-01.pdf")
# (3, 3)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-02.pdf")
# (3, 4)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-03.pdf")
# (4, 1)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-04.pdf")
# (4, 2)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-05.pdf")
# (4, 3)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-06.pdf")
# (4, 4)----------------------------------------------------
knitr::include_graphics("../docs/images/ecg-2021-05-07.pdf")
```

